---
title: "Pinniped Production"
output: pdf_document
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---


```{r,results="hide",echo=FALSE}
  suppressPackageStartupMessages(library(CIPinnipedAnalysis))
  source(file.path(system.file(package="CIPinnipedAnalysis"),"PupProduction.r"))
  maxyr=max(ZcProduction$Year)
```

# Zalophus 
```{r,results="hide",echo=FALSE,fig.height=7,fig.width=6,dpi=300}
par(mfrow=c(2,1))
plot(ZcProduction$Year[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],ZcProduction$TotalLiveCountByYear[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],xlab="Year",ylab="Live pup count",xaxt="n",main="Mainland SMI California sea lion Live Pup Count")
lines(ZcProduction$Year[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],ZcProduction$TotalLiveCountByYear[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"])
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))

zccr.years=ZcProduction$Year[ZcProduction$Area=="CastleRock"]
zccr.prod=ZcProduction$PupProduction[ZcProduction$Area=="CastleRock"]
zcyears=1970:max(zccr.years)
zcmissing.years=zcyears[!zcyears%in%zccr.years]
zccr.prod=c(zccr.prod,rep(NA,length(zcmissing.years)))
zccr.years=c(zccr.years,zcmissing.years)
zccr.prod=zccr.prod[order(zccr.years)]
zccr.years=sort(zccr.years)
plot(zccr.years,zccr.prod,main="Castle Rock California sea lion Live Pup Count",xlab="Year",ylab="Live pup count",xaxt="n",xlim=c(1970,maxyr))
lines(zccr.years,zccr.prod)
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))
dev.off()
```

Figure 1. Count of live Zalophus pups on San Miguel Island


```{r,results="hide",echo=FALSE,fig.width=6,fig.height=7,dpi=300}
plot(ZcProduction$Year[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],ZcProduction$MortalityRateAtLiveCount[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],xlab="Year",ylab="Mortality",xaxt="n",main="Zc SMI Mainland Pre-census Mortality")
lines(ZcProduction$Year[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],ZcProduction$MortalityRateAtLiveCount[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"])
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))
```

Figure 2. Pre-census Zalophus pup mortality rate


```{r,results="hide",echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
plot(ZcProduction$Year[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],ZcProduction$PupProduction[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],xlab="Year",ylab="Pup production",xaxt="n",main="SMI California sea lions Mainland")
lines(ZcProduction$Year[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"],ZcProduction$PupProduction[ZcProduction$Area=="Mainland"&ZcProduction$Island=="SMI"])
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2)) 

zccr.years=ZcProduction$Year[ZcProduction$Area=="CastleRock"]
zccr.prod=ZcProduction$PupProduction[ZcProduction$Area=="CastleRock"]
zcyears=1970:max(zccr.years)
zcmissing.years=zcyears[!zcyears%in%zccr.years]
zccr.prod=c(zccr.prod,rep(NA,length(zcmissing.years)))
zccr.years=c(zccr.years,zcmissing.years)
zccr.prod=zccr.prod[order(zccr.years)]
zccr.years=sort(zccr.years)
plot(zccr.years,zccr.prod,main="California sea lions Castle Rock",xlab="Year",ylab="Pup production",xaxt="n",xlim=c(1970,maxyr))
lines(zccr.years,zccr.prod)
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))

```

Figure 3. Estimated pup production of Zalophus pups on San Miguel Island


# Callorhinus 

```{r,results="hide",echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
plot(CuProduction$Year[CuProduction$Area=="Mainland"],CuProduction$TotalLiveCountByYear[CuProduction$Area=="Mainland"],xlab="Year",ylab="Live pup count",xaxt="n",main="Mainland Northern fur seal Live Pup Count",xlim=c(1970,maxyr),ylim=c(0,2500))
lines(CuProduction$Year[CuProduction$Area=="Mainland"],CuProduction$TotalLiveCountByYear[CuProduction$Area=="Mainland"])
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))
cr.years=CuProduction$Year[CuProduction$Area=="CastleRock"]
cr.prod=CuProduction$PupProduction[CuProduction$Area=="CastleRock"]
years=1970:max(cr.years)
missing.years=years[!years%in%cr.years]
cr.prod=c(cr.prod,rep(NA,length(missing.years)))
cr.years=c(cr.years,missing.years)
cr.prod=cr.prod[order(cr.years)]
cr.years=sort(cr.years)
plot(cr.years,cr.prod,main="Castle Rock Northern fur seal Live Pup Count",xlab="Year",ylab="Pup production",xaxt="n",xlim=c(1970,maxyr), ylim=c(0,1500))
lines(cr.years,cr.prod)
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))
```

Figure 4. Count of live Callorhinus pups on San Miguel Island

```{r,results="hide",echo=FALSE,fig.width=6,fig.height=7,dpi=300}
plot(CuProduction$Year[CuProduction$Area=="Mainland"],CuProduction$MortalityRateAtLiveCount[CuProduction$Area=="Mainland"],xlab="Year",ylab="Mortality",xaxt="n",main="Cu Mainland Pre-census Mortality",xlim=c(1970,maxyr))
lines(CuProduction$Year[CuProduction$Area=="Mainland"],CuProduction$MortalityRateAtLiveCount[CuProduction$Area=="Mainland"])
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))
```

Figure 5. Pre-census Callorhinus pup mortality rate


```{r,results="hide",echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
plot(CuProduction$Year[CuProduction$Area=="Mainland"],CuProduction$PupProduction[CuProduction$Area=="Mainland"],xlab="Year",ylab="Pup production",xaxt="n",main="Northern fur seals Mainland")
lines(CuProduction$Year[CuProduction$Area=="Mainland"],CuProduction$PupProduction[CuProduction$Area=="Mainland"])
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))

cr.years=CuProduction$Year[CuProduction$Area=="CastleRock"]
cr.prod=CuProduction$PupProduction[CuProduction$Area=="CastleRock"]
years=1970:max(cr.years)
missing.years=years[!years%in%cr.years]
cr.prod=c(cr.prod,rep(NA,length(missing.years)))
cr.years=c(cr.years,missing.years)
cr.prod=cr.prod[order(cr.years)]
cr.years=sort(cr.years)
plot(cr.years,cr.prod,main="Northern fur seals Castle Rock",xlab="Year",ylab="Pup production",xaxt="n",xlim=c(1970,maxyr))
lines(cr.years,cr.prod)
axis(1,at=seq(1970,maxyr,2),labels= seq(1970,maxyr,2))
```

Figure 6. Estimated pup production of Callorhinus pups on San Miguel Island




