---
title: "Zalophus pup weights"
output: pdf_document
date: '`r format(Sys.Date(), "%B %d, %Y")`'
---


```{r,results="hide",echo=FALSE}
  suppressPackageStartupMessages(library(CIPinnipedAnalysis))
#  doc.type <-knitr::opts_knit$get('rmarkdown.pandoc.to')
  lastyear=2014
  sdir=system.file(package="CIPinnipedAnalysis")
  source(file.path(sdir,"ZC_Weight_Adjustment_Model.r"))
```


```{r,echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
maxyear=max(zcweights$cohort)
ymin=min(c(female.averages$fit-1.96*female.averages$se, male.averages$fit-1.96*male.averages$se))*.9
ymax=max(c(female.averages$fit+1.96*female.averages$se, male.averages$fit+1.96*male.averages$se))*1.1
plot_weight.series(1975:maxyear,female.averages,main="California sea lion Female Pups",ylim=c(ymin,ymax),date="1 Oct")
plot_weight.series(1975:maxyear,male.averages,main="California sea lion Male Pups",ylim=c(ymin,ymax),date="1 Oct")
```

Figure 1. Predicted Zalophus pup weights on San Miguel Island at 1 Oct. 


```{r,echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
maxyear=max(zcweights$cohort)
female.mean=mean(female.averages$fit)
male.mean=mean(male.averages$fit)
ymin=min(c(female.averages$fit-1.96*female.averages$se, male.averages$fit-1.96*male.averages$se))*.9
ymax=max(c(female.averages$fit+1.96*female.averages$se, male.averages$fit+1.96*male.averages$se))*1.1
female.averages$fit=female.averages$fit-female.mean
plot_weight.series(1975:maxyear,female.averages,main="California sea lion Female Pups",ylim=c(ymin-female.mean,ymax-female.mean),date="1 Oct")
abline(h=0)
male.averages$fit=male.averages$fit-male.mean
plot_weight.series(1975:maxyear,male.averages,main="California sea lion Male Pups",ylim=c(ymin-male.mean,ymax-male.mean),date="1 Oct")
abline(h=0)
```

Figure 2. Predicted Zalophus pup weights on San Miguel Island at 1 Oct expressed as anomaly from overall mean. 


```{r,echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
maxyear=max(zcweights$cohort)
ymin=min(c(winter.female.averages$fit-1.96*winter.female.averages$se, winter.male.averages$fit-1.96*winter.male.averages$se))*.9
ymax=max(c(winter.female.averages$fit+1.96*winter.female.averages$se, winter.male.averages$fit+1.96*winter.male.averages$se))*1.1
plot_weight.series(1975:maxyear,winter.female.averages,main="California sea lion Female Pups",ylim=c(ymin,ymax),date="1 Feb")
plot_weight.series(1975:maxyear,winter.male.averages,main="California sea lion Male Pups",ylim=c(ymin,ymax),date="1 Feb")
```

Figure 3. Predicted Zalophus pup weights on San Miguel Island at 1 Feb. 

```{r,echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
maxyear=max(zcweights$cohort)
female.mean=mean(winter.female.averages$fit)
male.mean=mean(winter.male.averages$fit)
ymin=min(c(winter.female.averages$fit-female.mean-1.96*winter.female.averages$se, winter.male.averages$fit-male.mean-1.96*winter.male.averages$se))*.9
ymax=max(c(winter.female.averages$fit-female.mean+1.96*winter.female.averages$se, winter.male.averages$fit-male.mean+1.96*winter.male.averages$se))*1.1
plot_weight.series(1975:maxyear,winter.female.averages-female.mean,main="California sea lion Female Pups",ylim=c(ymin,ymax),date="1 Feb")
abline(h=0)
plot_weight.series(1975:maxyear,winter.male.averages-male.mean,main="California sea lion Male Pups",ylim=c(ymin,ymax),date="1 Feb")
abline(h=0)
```

Figure 4. Predicted Zalophus pup weights on San Miguel Island at 1 Feb expressed as anomaly from overall mean. 


```{r,echo=FALSE,fig.width=6,fig.height=7,dpi=300}
par(mfrow=c(2,1))
with(ZCWeight.df,plot(female.observed.mean.fall[fall.avg.day< -7],female.adjusted.mean.fall[fall.avg.day< -7],col="blue",ylim=c(12,23),xlim=c(12,23),xlab="Fall observed mean",ylab="Fall adjusted mean"))
with(ZCWeight.df,points(female.observed.mean.fall[fall.avg.day >= 7],female.adjusted.mean.fall[fall.avg.day>= 7],col="red"))
with(ZCWeight.df,points(female.observed.mean.fall[fall.avg.day >= -7 &fall.avg.day< 7],female.adjusted.mean.fall[fall.avg.day>= -7 &fall.avg.day<7],col="black"))
abline(0,1)
# show plot of observed and predicted from adjustment mean for fall
with(ZCWeight.df,plot(female.observed.mean.winter[winter.avg.day< 116],female.adjusted.mean.winter[winter.avg.day< 116],col="blue",ylim=c(14,32),xlim=c(14,32),xlab="Winter observed mean",ylab="Winter adjusted mean"))
with(ZCWeight.df,points(female.observed.mean.winter[winter.avg.day >= 130],female.adjusted.mean.winter[winter.avg.day>= 130],col="red"))
with(ZCWeight.df,points(female.observed.mean.winter[winter.avg.day >= 116 &winter.avg.day< 130],female.adjusted.mean.winter[winter.avg.day>= 116 &winter.avg.day< 130],col="black"))
abline(0,1)
```

Figure 5. Predicted and observed Zalophus pup weights on San Miguel Island at 1 Oct (Fall) and 1 Feb (winter) using adjustment model. Blue values are 7 days or after the standardized date. Red values are less than 7 days before the standardized date.  Black values are within 7 days of the standardized date.

